\documentclass[letterpaper,10pt,headsepline]{scrreprt}
\usepackage[latin1]{inputenc}
\usepackage[ngerman]{babel}
\usepackage[T1]{fontenc}
\usepackage[dvips]{graphicx}
\usepackage{mathpazo}
\usepackage{microtype}
\usepackage{scrpage2}
\usepackage{paralist}
\clubpenalty=6000
\widowpenalty=6000
\author{Ross Laird}
\title{Creative Writing 1240}
\date{today}
\ohead{Creative Writing 1240}
\pagestyle{scrheadings}
\setcounter{secnumdepth}{-1}
\begin{document}
\section{Creative Writing 1240}
Instructor: Ross Laird, Ph.D.\\
Telephone: 604-916-1675\\
Email: ross@rosslaird.com\\
Website: www.rosslaird.com\\
Location: Kwantlen Polytechnic University, Surrey Campus\\
Schedule: Thursdays, 11:00-12:50\\
January 8 to April 8, 2009\\

\subsection{Basic Philosophy of the Course}

This course continues the foundational work begun in Interdisciplinary Expressive Arts 3100 or Mythological Narratives 3130, in which students have explored the many dimensions of interdisciplinary inquiry. The course is, for the most part, led by the initiatives and interests of students working individually and in groups. The course content will evolve as the course develops, but will include a diverse number of current interdisciplinary approaches within the expressive arts. These include creative writing, music, movement, fine arts, theatre, expressive arts therapies, storytelling, mythopoetics, photography, film making, etc.. Additionally, students will explore potential academic and career directions within the cultural and social services sectors.

As a reminder, and for new students (those who have not enrolled in IDEA 3100), Interdisciplinary Expressive Arts refers to a specific set of educational activities, goals and strategies. Based on innovative
pedagogy and integrative approaches to learning, interdisciplinary
studies involve the synthesis and synergy of various disciplines
toward a cohesive, unified educational experience. Interdisciplinarity
is much more than enrollment in courses from more than a single
discipline. Authentic interdisciplinarity emphasizes the linkages
between disciplines by focusing on contrasting and complementary
aspects of diverse educational domains.

Interdisciplinary studies encourage students to develop broader
intellectual skills, greater facility for critical thinking, and
greater awareness of the social relevance of their education.
Interdisciplinary students have the opportunity to develop exemplary
skills in problem solving, insight, team-building, lateral thinking,
and multi-modal learning styles. Interdisciplinary strategies involve
approaching an issue or problem from various perspectives. This
typically entails intellectual inquiries that range beyond the borders
of any single discipline or domain. While still respecting the
function of the boundaries between domains, interdisciplinary
approaches recognize that those boundaries are essentially arbitrary
and do not always serve the goals of learning. Global warming and the
AIDS pandemic are two examples of contemporary issues that require
interdisciplinary approaches.

This course is about creativity, about making a claim for the
fundamental right of intentional creative action. Within that context,
we will explore the ancient and modern practices of creative endeavor
(particularly as regards family and culture), the hurdles of
creativity (as they involve craft and precision and clarity) and the
great gifts we might receive from others of our creative kin (that is
to say, the long tradition of writers, poets, sculptors, dancers,
craftspeople of all stripes, musicians, myth-makers, and so on).
Throughout this process, our guiding archetype will be that of the
trickster.

In this course we stake out the territory of the creative, inspecting
the geology of its forms and ideals, finding our own individual places
to homestead. Creativity involves the search for truth, yet also an
awareness that truth and fact are often provisional, and mythological;
they are shapeshifters on the wide-open plain of creativity. We will
explore what this means, and what to do about it.

And, finally, the goal of the course (from my point of view, at
least), is to have fun: to preserve and nurture the creative and
imaginative spirit that is the foundation of all the arts and
sciences.

\subsection{Learning Goals}
\begin{itemize}

\item Read selected interdisciplinary texts and discuss their origin, development, and contemporary relevance
\item Interpret interdisciplinary literary traditions within the context of contemporary academic and professional inquiry
\item Articulate (verbally, in writing, and through creative endeavour) knowledge of interdisciplinarity as both an ancient and a current mode of inquiry
\item Describe the ways in which interdisciplinary expressive arts are more than a simple aggregation of various disciplines
\item Evaluate diverse interdisciplinary perspectives and approaches using heuristic modalities
\item Complete a series of linked interdisciplinary projects and presentations
\item Complete a final project using interdisciplinary principles

\end{itemize}
\clearpage

\subsection{Learning Experiences}
The course will include a variety of learning experiences contingent upon regular attendance and dedicated participation. Because creativity is an interactive process, much of the class time will be devoted to group experiential exercises, individual reflective tasks, collaborative endeavors, and practical assignments.

We will create a collaborative environment in this class. We are not going to cobble together the type of group one often hears about in the arts: competitive, cut-throat, critical. Repeat: we are not creating such a group. Instead, we will direct our efforts toward building upon the individual strengths of each participant, finding ways for each of us to be self-reflective in terms of assessing our creative work, discovering a means of protecting the quality and integrity of our writing. The creative spirit is remarkably persistent, yet it is also fragile, especially at its inception, and
we must be conscious of this fragility. Think about it: did you not experience, as a child, the strangulation of your creativity in school, by way of a culture of insensitive peers or teachers? Why do you think hardly anyone feels comfortable singing in public, or dancing, or drawing, or reading their written work to others? We have, most of us, been the victims of inappropriate feedback and judgment. We have to be careful about this, in our course, so that we do not harm one another.

\section{Readings}
\subsection{Required Course Texts}

\begin{description}
\item [Barron, F., Montouri, A., and Barron, A., eds.] \textit{Creators on Creating: Awakening and Cultivating the Imaginative Mind.}
\\New York: Putnam, 1997.
\item[Wilson, Frank.] \textit{The Hand: How Its Use Shapes the Brain, Language and Human Culture.}
\\New York: Vintage, 1998.
\item[Plus] {one text chosen by the student.}
\end{description}

\subsection{Suggested Books}
\begin{description}
\item [Allen, Pat.] \textit{Art is a Way of Knowing.} \\Shambhala, 1995.
\item [Hyde, Lewis.] \textit{The Gift: Imagination and the Erotic Life of Property.}
\\New York: Vintage, 1983.
\item [Butala, Sharon.] \textit{Wild Stone Heart}. \\HarperFestival,
  2000. \textsc{ISBN 000255397X}.
\item [Calvo, C\'esar.] \textit{The Three Halves of Ino Moxo}.
  \\Translated by Kenneth Symington. \\Inner Traditions, 1995.
  \textsc{ISBN 0892815191}.
\item [Campbell, Joseph.] \textit{The Mythic Image}.
  \\Princeton UP, 1974.
\item [Ellis, Normandi.] \textit{Dreams of Isis: A Woman's Spiritual
    Sojourn}.
  \\Quest, 1995.
\item [Hancock, Graham.] \textit{Heaven's Mirror: Quest for the Lost
    Civilization.}.
  \\Crown, 1998.
\item [Hedges, Chris.] \textit{War Is a Force that Gives Us Meaning}.
  \\Anchor, 2003. \textsc{ISBN 1400034639}.
\item [Kingston, Maxine Hong.] \textit{The Woman Warrior: Memoirs of a
    Girlhood Among Ghosts}. \\Vintage, 1989. \textsc{ISBN
    0072435194}.
\item [Kwan, Michael David.] \textit{Things that Must Not be
    Forgotten: A Childhood in Wartime China}. \\Soho Press
  \textsc{ISBN 1569472823}
\item [Laird, Ross A.] \textit{Grain of Truth: The Ancient Lessons of Craft}. \\MWR, 2000.
\item [Langewiesche, William.] \textit{American Ground: Unbuilding
    \\the World Trade Center}. \\North Point Press, 2002.
  \textsc{ISBN 0865475822}. (Also see \textit{Inside the Sky}.)
\item [London, Peter.] \textit{No More Secondhand Art.}
\\Boston: Shambhala, 1989.
\item [Lopate, Phillip.] \textit{The Art of the Personal Essay: An
    Anthology from the Classical Era to the Present}. \\Anchor, 1997.
  \textsc{ISBN 038542339X}.
\item [Macfarlane, David.] \textit{The Danger Tree: Memory, War and
    the Search for a Family's Past}. \\Walker, 2001. \textsc{ISBN
    0802776167}.
\item[McNiff, S.] \textit{Art Heals: How Creativity Cures the Soul.} \\Boston: Shambhala Publications, 2004.
\item [Merwin, W.S.] \textit{The Mays of Ventadorn}. \\National
  Geographic Directions, 2002. \textsc{ISBN 0792265386}.
\item [Ondaatje, Michael.] \textit{Running in the Family}. \\Vintage,
  1993. \textsc{ISBN 0679746692}.
\item [Pirsig, Robert.] \textit{Zen and the Art of Motorcycle
    Maintenance}. \\HarperTorch, 2006 (reprint). \textsc{ISBN
    0060589469}.
\item [Saint-Exup\'ery, A.] \textit{Wind, Sand and Stars}. \\Harvest,
  2002. \textsc{ISBN 0156027496}.
\item [Sanders, Scott Russell.] \textit{Writing from the Center}.
  \\Indiana UP, 1997. \textsc{ISBN 0253211433}.
\item [Sullivan, William.] \textit{The Secret of the Incas: Myth,
    Astronomy, and the War Against Time.} \\Crown, 1996.

\end{description}

\subsection{Books on Creativity and Associated Philosophies}
\begin{description}
\item [Achebe, Chinua] \textit{Hopes and Impediments}. New York:
  Doubleday, 1989.
\item [Barron, F., ed] \textit{Creators on Creating: Awakening and
    Cultivating the Imaginative Mind}. New York: Putnam, 1997.
\item [Benjamin, Walter] \textit{Theses on the Philosophy of History}.
\item [Borges, Jorge Luis.] \textit{Collected Fictions}. \\Penguin,
  1999. \textsc{ISBN 0140286802}.
\item [Bohm, David] \textit{Wholeness and the Implicate Order}.
  London: Ark, 1980.
\item [Bohm, David] \textit{Unfolding Meaning}. New York: Routledge,
  1985.
\item [Bohm, David] \textit{On Creativity}. New York: Routledge, 1998.
\item [Bronowski, Jacob] \textit{Science and Human Values}. New York:
  Harper, 1956.
\item [---------] \textit{The Face of Violence}. London: Turnstile
  Press, 1964.
\item [---------] \textit{A Sense of the Future: Essays in Natural
    Philosophy}. Cambridge, MIT Press, 1977.
\item [Degler, Teri] \textit{The Fiery Muse: Creativity and the
    Spiritual Quest}. Toronto: Random House, 1996.
\item [Demos, John.] \textit{The Unredeemed Captive: A Family Story
    from Early America}. \\Vintage, 1995. \textsc{ISBN 0679759611}.
\item [Flack, Audrey] \textit{Art and Soul: Notes on Creating}. New
  York: Penguin, 1986.
\item [Franklin, Ursula] \textit{The Real World of Technology}.
  Toronto: Anansi, 1999.
\item [Fulford, Robert] \textit{The Triumph of Narrative: Storytelling
    in an Age of Mass Culture}. Toronto: Anansi, 1999.
\item [Goldberg, Natalie] \textit{Writing Down the Bones}. Boston:
  Shambhala, 1986.
\item [Herrigel, Eugen] \textit{Zen in the Art of Archery}. New York:
  Random House, 1977.
\item [Hildegard of Bingen] \textit{Secrets of God: Writings of
    Hildegard of Bingen}.\\ Boston: Shambhala, 1996.
\item [Hyde, Lewis] \textit{The Gift: Imagination and the Erotic Life
    of Property}. New York: Vintage, 1983.
\item [---------] \textit{Trickster Makes This World: Mischief, Myth,
    and Art}. New York: North Point Press, 1998.
\item [Jim\'enez, Juan Ramon] \textit{The Complete Perfectionist: A
    Poetics of Work}. Edited and translated by Christopher Maurer. New
  York: Doubleday, 1997.
\item [Jung, C.G] \textit{The Spirit in Man, Art and Literature}.
  Translated by R.F.C. Hull. Princeton: Princeton University Press,
  1998.
\item [London, Peter] \textit{No More Secondhand Art}. Boston:
  Shambhala, 1989.
\item [Pye, David] \textit{The Nature and Art of Workmanship}.
  Cambridge: Cambridge UP, 1968.
\item [Richards, Mary] \textit{Centering in Pottery, Poetry and the
    Person}. Middletwon, CT: Wesleyan UP.
\item [Sarton, May] \textit{Journal of a Solitude}. New York: Norton,
  1973.
\item [Sennett, Richard] \textit{The Corrosion of Character: The
    Personal Consequences of Work in the New Capitalism}. New York:
  Norton, 1998.
\item [Thoreau, Henry David] \textit{Walden}. New York: Norton, 1985.
\end{description}
\newpage

\subsection{Further Suggestions}

\begin{description}

\item [Arendt, Hannah] \textit{Illuminations}. London: Cape.

\item [Chodorow, J.]  \textit{Dance Therapy and Depth Psychology: The Moving Imagination}. Routledge, 1991.

\item [Dewey, John] \textit{Art As Experience}. New York, Perigee, 1931.
\item [Diamonstein, Barbara] \textit{Handmade in America: Conversations with Fourteen Craftmasters}. New York: Abrams, 1983.
\item [Gadamer, H.G.] \textit{Philosophical Hermeneutics}. Trans. D.E. Linge. Berkeley: University of California Press, 1976.
\item [Gardner, Harold] \textit{The Unschooled Mind: How Children are Taught and How Teachers Should Teach}. New York, Basic, 1993.
\item [Greene, Brian] \textit{The Elegant Universe: Superstrings, Hidden Dimensions, and the Quest for the Ultimate Theory}. New York: Norton, 1999.
\item [Hammarskjold, Dag] \textit{Markings}.
\item [Knill, P., Levine, E. and Levine, S.] \textit{Principles and Practice of Expressive Arts Therapy}. New York: Jessica Kingsley, 2004.
\item [Levine, S. K. and Levine, E. G.] \textit{Foundations of Expressive Arts Therapy}. London: Jessica Kingsley Publishers, 1999.
\item [Levine, E.] \textit{Tending the Fire}. Ontario: EGS Press, 2003.
\item [Levine, Peter] \textit{Waking the Tiger: Healing Trauma}. North Atlantic Books.
\item [Lorca, Federico] \textit{In Search of Duende}. Translated by Christopher Maurer. New York: New Directions, 1998.
\item [Lyndon, Susan] \textit{The Knitting Sutra: Craft as a Spiritual Practice}. San Francisco: Harper, 1997.
\item [McNiff, Shaun] \textit{Art Heals: How Creativity Cures the Soul}. Boston: Shambhala Publications, 2004.
\item ---------. \textit{Creating with Others: The Practice of Imagination in Art, Life, and the Workplace}. Boston: Shambhala Publications, 2003.
\item ---------. \textit{Art As Medicine}. Boston: Shambala, 1992.
\item ---------. \textit{Fundamentals of Art Therapy}. Springfield, IL: Charles C. Thomas, 1988.
\item ---------. \textit{Art-Based Research}. London: Jessica Kingsley, 1988.
\item [Mazza, Nick] \textit{Poetry Therapy: Interface of the Arts and Psychology}. CRC Press, 1999.
\item [Minnich, E. K.] \textit{Transforming Knowledge}. Temple University Press., 1990.
\item [Moustakas, C.] \textit{Heuristic Research: Design, Methodology, and Applications}. Newbury Park: Sage, 1990.
\item [Needleman, Carla] \textit{The Work of Craft: An Inquiry Into the Nature of Crafts and Craftsmanship}. New York: Kodansha, 1979.

\end{description}

\clearpage

\section{Demonstration of Learning}

\subsection{Assignments}
Three linked assignments and several presentations (see below) are required for this course. These assignments may be comprised of any type of art expression (writing, music, imagery, dance, movement, photography, etc.). The central idea is for you to choose a specific theme or thread and explore it in some depth. We will discuss these projects at length in class. They are opportunities for you to discover and explore creativity in your own life.

For philosophical reasons, I do not prescribe a particular length or structure for the projects. There is no upper limit on the length or complexity of the projects.

The individual projects are worth 25 percent each.

\newpage

\subsection{Group Presentations}

Each student will be a member of several different peer groups; each
peer group will present at least one mini-presentation (roughly thirty minutes each) on various topics. Each class session after the second will involve presentations, with one presentation from each group. Class time will be given for preparing the presentations. The structure and content of the presentations will be discussed in class.

The group presentations are worth a total of 25 per cent of your grade.

\subsubsection{Presentation Methods and Goals}

The central idea of the presentations for this course is to give you
 opportunities to practice interdisciplinary thinking and expression.
 As such, the presentation should be interdisciplinary. Essentially,
 this means that you should try to use multiple presentation
 strategies and modalities. These might include (but are certainly not
 limited to) any of the following:

 \begin{itemize}
 \item Storytelling
 \item Poetry
 \item Music (playing)
 \item Drumming
 \item Singing
 \item Dance
 \item Movement
 \item Sport
 \item Ritual
 \item Film (showing)
 \item Film making
 \item Photography
 \item Web content
 \item Craft work
 \item Art making
 \item Individual reflection
 \item Meditation
 \item Health practices
 \item Creative process (any type)
 \item Group communication
 \item Cultural practices
 \item Nature experiences
 \end{itemize}

 Whenever possible (and workable), try to mix together multiple
 modalities into a single presentation. For example, you might ask the
 group to do some individual reflection using the modality of poetry,
 then create a series of movements based on the poetry, then work in
 small groups to talk about and share the process. Many configurations
 are possible. The trick is to choose an activity that you enjoy, then
 find a way to apply it to the content (suggested presentation topics
 are listed below). Please do not create your presentations using only
 written and/or spoken materials. In other words, don't just stand up
 at the front of the class and talk about the presentation topic.
 Utilize the energy of the group. Remember that in interdisciplinary
 work divergences are valued as unique opportunities. So, feel free to
 experiment with activities and modalities that may not seem, on the
 surface, to be related to the topic at hand but which might, upon
 experiment, yield surprising connections and results. Be playful.
 Allow yourself to laugh at yourself, to be embarrassed, to engage
 with the process in novel and interesting ways.

In interdisciplinary work, riddles and puzzles are highly prized.
Accordingly, the presentations should (ideally) not be complete
explanations or presentations of material. Feel free to play with
challenging exercises, with impossible scenarios, and other conundra.
One way to think about this is to consider insoluble riddles, such as
the one in \textit{Alice in Wonderland}: Why is a raven like a
writing desk?

\begin{quote}
  ``Have you guessed the riddle yet?'' the Hatter said, turning to Alice again.\\
``No, I give it up,'' Alice replied. ``What's the answer?''\\
``I haven't the slightest idea,'' said the Hatter.\\
``Nor I,'' said the March Hare.\\
Alice sighed wearily. ``I think you might do something better with the time,'' she said, ``than wasting it in asking riddles that have no answers.''
\end{quote}

The best interdisciplinary topics offer more questions than answers.
They, are essentially, gateways into the mysterious--which, as
Einstein will tell you, is an important place to be:

\begin{quote}
  The most beautiful thing we can experience is the mysterious. It is
  the source of all true art and science.
\end{quote}

\subsubsection{Suggested Topics for Interdisciplinary Presentations}

\begin{itemize}
\item Akhenaten and the invention of monotheism
\item Albrecht Durer and alchemy
\item Aristotle's book of comedy
\item Bill Evans and the Peace Piece
\item Buckminster Fuller and the geodesic
\item Chenrizi and the politics of China
\item Chuang Tzu and the butterfly
\item Coleridge and the person from Porlock
\item Csikszentmihalyi and the flow experience
\item David Bohm's Implicate Order
\item Darwin, the bassoon, and the sundew
\item Eugen Herrigel and the practice of archery
\item Francis Yates and the \textit{Art of Memory}
\item Freud, Jung, and the ``bosh'' incident
\item Fulcanelli and \textit{Mysteries of the Cathedrals}
\item Giordano Bruno and the Hermetic tradition
\item Godel's uncertainty principle
\item Hanna Arendt at Nuremberg
\item Henri Rousseau in the jungle
\item Howard Carter and ``wonderful things''
\item Jacob Bronowski at Auschwitz
\item Jacob Bronowski, Nagasaki, and \textit{Science and Human Values}
\item Jan Tschichold and the Nazis
\item John Cage on the subway with the \textit{I Ching}
\item Kepler's \textit{Somnium}
\item Mary Shelley and the genesis of \textit{Frankenstein}
\item Newton's \textit{Principia}
\item Nikola Tesla and universal energy
\item Philip K. Dick, VALIS, and 2-3-74
\item Picasso, Guernica, and Expo 1937
\item R.D. Laing and madness as reality
\item Ramanujan's notebooks
\item Richard Feynman and the invention of quantum mechanics
\item Schwaller de Lubicz at Karnak
\item Simone Weil and leading from desire
\item St. Exupery flying into the desert
\item The Reimann Hypothesis
\item The Voynich Manuscript
\item The visions of Hildegard of Bingen
\item Thoth's legacy
\item Walter Benajmin and the Angel of History
\item Wendell Berry going \textit{Into the Woods}
\item Wilhelm Reich's Cloudbuster
\item William Blake's \textit{Marriage of Heaven and Hell}

\end{itemize}


\subsection{Evaluation of Assignments and Presentations}
My primary focus, as an instructor, is to assist you in developing
your creativity. Grades are quite far down on the list of priorities
for me. I am focused on your engagement with the process, your
commitment to your own work, the extent to which you show up,
metaphorically, to be as present as you can be. These are evaluation
criteria for me.

\subsection{Attendance and Participation}
The expectation is that you will attend all sessions and involve
yourself in the class process. Your willingness to engage creatively
with the learning process, to take appropriate personal risks, and to
participate in group activities are all central to your involvement in
this class. Your emotional involvement in the class is
as important as your academic knowledge of the material.

\subsection{Grade Inflation}
Every semester there are students who do well on the
assignments, complete all the associated learning goals of the course,
participate well, and wonder why they do not receive a grade of one
hundred percent (or 98, anyway). Here is the reason: almost every
semester there are students who demonstrates a level of commitment
that goes beyond the course requirement. Such students complete extra
work, or hand in exemplary assignments, or undertake a significant
amount of personal development in addition to the course expectations.
Such students typically receive the highest grades.

If you do reasonably well in the course you will receive a reasonable
grade. Very high grades are intended for extra or exemplary work.
Unfortunately, over the past thirty years the post-secondary
educational system in North America has participated in a process of
grade inflation. Since the 1980's, the average grade for typical
course work has been increasing by about 25 per cent each decade.
Elevated assessments do not accurately reflect the work of most
students. Even worse, grade inflation has caused many students to
expect high grades for average work. I am not a particularly stringent
assessor; but I will not inflate grades artificially.

The grades for the course will be distributed along a curve, with a
small number of students (likely) receiving high grades, most students
receiving grades in the middle range, and a few students struggling
with lower grades. If you are uncertain about your assessment for a
given assignment, or if you wish to know where, roughly, you are along
the distribution curve of the class, or if you would like suggestions
for how to improve your grade, please ask me for clarification.
\clearpage

% \section{Class Schedule}
% The class sessions will be balanced between presentations (by the % instructor and students) academic material, group collaboration, and % composition. The content for each session will evolve as the semester % progresses. We will cover the following themes (though, perhaps not in % the order listed below):
% \\
% \begin{compactdesc}

% \item[The Nature of Interdisciplinarity]
% A vignette: Bronowski at Nagasaki.
% Definitions and principles.
% Clarifications of common misunderstandings about interdisciplinary % approaches and practices.
% An introduction to interdisciplinarity as a mode of inquiry in the % arts and sciences.
% \\
% \item[Traditions and Practices]
% Consideration of interdisciplinary practices as effective vehicles for % the transmission of  sacred, social, political, artistic, and % scientific information.
% Examination of interdisciplinarity as a fundamental and necessary % function of human nature and inquiry.
% \\
% \item[Developments and Milestones]
% Consideration of the development of interdisciplinarity and its role % in the contemporary world.
% Examination of the relationship between interdisciplinary practices % and traditional, faculty-based inquiry in the academic environment.
% \\
% \item[Themes and Philosophies]
% Introduction to the historical background of interdisciplinary % philosophy and practice in science and literature.
% Explication of ancient world views, with particular emphasis on % spirituality, science, mythological concepts, and approaches to the % imagination.
% \\
% \item[Hermetic Threads]
% Reading of excerpts from core Egyptian, alchemical, and early % scientific texts, with particular emphasis on foundational % interdisciplinary ideas.
% Examination of the ways in which the sciences and the arts were % entwined in the practices and perspectives of all peoples until the % twentieth century.
% Exploration of the transmission of interdisciplinary ideas into the % contemporary world.
% \\
% \item[Interdisciplinarity in the Expressive Arts]
% Exploration of the transmission of interdisciplinary philosophies and % practices by way of the expressive arts.
% Examination of practices and structures within the interdisciplinary % expressive arts (writing, storytelling, dance, movement, ritual, % religious practice, music, philosophy, etc.), with emphasis on the % traditions of psychology and mythology.
% Reading of selected texts within the interdisciplinary expressive % arts.
% Examination of the ways in which texts within the interdisciplinary % expressive arts have influenced the development of the arts and % sciences.
% \\
% \item[Multicultural Principles]
% Introduction to the interdisciplinary contributions made by Asian % mythology, literature, and science.
% Explication of ancient Asian world views, with particular emphasis on % interdisciplinary spirituality, mythological concepts, and approaches % to knowledge.
% Reading of excerpts from Asian texts, with particular emphasis on the % synthesis of interdisciplinary expressive arts within those % traditions.
% \\
% \item[Other Stuff (that we'll think up as we go along)]
% Much of the course content will be generated by students.

% \end{compactdesc}



\end{document}
