   \TeXXeTstate=1
\documentclass[letterpaper,10pt,DIV=9,headsepline]{scrreprt}
   \usepackage{fontspec,xunicode}
    \defaultfontfeatures{Numbers=OldStyle,Scale=MatchLowercase,Mapping=tex-text}
    \setmainfont{Adobe Jenson Pro}
    \setromanfont[Mapping=tex-text]{Adobe Jenson Pro}
    \setsansfont[Scale=MatchUppercase]{Brioso Pro}
   \usepackage{graphicx}
   \usepackage{xltxtra}

\usepackage{scrpage2}
\clubpenalty=6000
\widowpenalty=6000
\author{Ross Laird}
\title{IDEA 3100 Assessment Criteria}
\date{today}
\ohead{IDEA 3100}
\begin{document}
\pagestyle{scrheadings}
\setcounter{secnumdepth}{-1}

\subsubsection{Assessment Criteria for IDEA 3100}

\begin{itemize}
\item Demonstration of commitment to interdisciplinary process.
\item Demonstration of commitment to the development of self-awareness.
\item Demonstration of commitment to the development of creativity.
\item Openness to interpersonal and collaborative process.
\item Ability to participate in appropriate creative exploration.
\item Consideration of and responsiveness to others.
\item Willingness to take appropriate risks and to challenge oneself.
\item Commitment to enhancing the interpersonal experience of everyone
  in the class.
\item Commitment to enhancing the interdisciplinary experience of
  everyone in the class.
\item Ability to take personal responsibility for learning.
\item Willingness to deal with conflicts appropriately if and when they arise.
\item Ability to be open and responsive to appropriate feedback.
\item Willingness to speak up, to join conversations, and to contribute.
\end{itemize}

\end{document}
%%% Local Variables: %%% mode: latex %%% TeX-master: t %%% End:
