%!TEX encoding = UTF-8 Unicode
%    \documentclass[10pt, letterpaper]{article}
\documentclass[letterpaper,10pt,headsepline]{scrreprt}
    \usepackage{fontspec}
    \usepackage{placeins}
    \usepackage{color}
    \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
    \usepackage{multibbl}
    \usepackage{graphicx}
    \usepackage{hieroglf}
    \usepackage{marginnote}
    \usepackage{mparhack}
    \usepackage{txfonts}
    \usepackage{url}
    \usepackage{listings}
    \lstset{language=HTML}
    \usepackage{titling}
    \usepackage{geometry}
    \geometry{letterpaper, textwidth=5.5in, textheight=8.5in, marginparsep=7pt, marginparwidth=.6in}
    %\setlength\parindent{0in}
    \defaultfontfeatures{Mapping=tex-text}
    \setromanfont [Ligatures={Common}, SmallCapsFont={Sabon LT Std}, BoldFont={Sabon LT Std Bold}, ItalicFont={Sabon LT Std Italic}]{Sabon LT Std}
    \setmonofont[Scale=0.8]{Lucida Sans Typewriter Std}
    \setsansfont [Ligatures={Common}, SmallCapsFont={Myriad Pro Regular}, BoldFont={Myriad Pro Bold}, ItalicFont={Myriad Pro Italic}]{Myriad Pro Italic}
\usepackage[ngerman,english]{babel}
\usepackage{scrpage2}
\usepackage{paralist}
\clubpenalty=6000
\widowpenalty=6000
\author{Ross Laird and Jody Gordon}
\title{Katalyst Grant Application}
\date{\today}
\ohead{Katalyst}
\chead{Research Grant Application}
\pagestyle{scrheadings}
\setcounter{secnumdepth}{-1}
%\contentsname{Contents}
\begin{document}
\begin{titlingpage}
\begin{center}
\maketitle
\end{center}
\end{titlingpage}
\setcounter{tocdepth}{3}
\tableofcontents

\vspace*{1cm}

\section{Proposed Budget}

\subsection{Time Release}

We are requesting two semesters of one-quarter time release for Ross Laird to
work on this project and to write the first draft of the book that will
represent the broad narrative component of our research results. This includes
planning, outlining, synthesizing and interpreting the heuristic data, writing
various drafts, fact-checking, editing, copy-editing, designing, typesetting,
digitizing (in multiple formats), and arranging for publication. Writing a
book is a substantial effort and requires a significant investment in time and
focus.

\$25,000 is the amount allocated to two sections of time release, but it is
also the amount specified by the Canada Council for grants to writers in the
category to which Ross Laird belongs: mid-career writers who have published
two or more books. Canada Council grants for writers are intended to provide
``Canadian authors (emerging, mid-career and established) time to write new
literary works.'' This is precisely what Ross Laird will be doing, and the Canada
Council's industry-standard is the best guide to budgetary justification for
this type of work.

\subsection{Funding for Student Researchers}

We are requesting a total of \$10,000 in funding to support the collaborative
work of ten students. These students will be partners in our research and will
be fully involved in all of our activities. We project that the activities of
the students will require roughly 65 hours in total, and we propose to pay the
students \$15 per hour.

\subsection{Funding for Attendance at Conferences}

We request a total of \$5,000 to support the travel of the two research leads
to the First-Year Experience Conference, which is the largest and most
multidisciplinary (as well as interdisciplinary) conference on the first year
experience in the United States. This conference will be held in February
2013, and the work presented will be directly relevant to our research at
Kwantlen. Attendance at this conference by the two research leads (with the
added possibility of presenting our work to date at this conference) will
greatly inform us about high-impact practices occurring at universities and
colleges across North America.

\subsection{Total Budget}

The total projected cost of the project is \$40,000.

\section{Timeline}

The timeline for this project is two years.


\end{document}
