\TeXXeTstate=1
\documentclass[10pt,DIVcalc,oneside,letterpaper,headexclude]{scrreprt}
\usepackage{fontspec,xunicode}
\defaultfontfeatures{Numbers=OldStyle,Scale=MatchLowercase,Mapping=tex-text}
\setmainfont{Sabon LT Std}
\setromanfont[Mapping=tex-text]{Sabon LT Std}
\setsansfont[Scale=MatchUppercase,Letters=SmallCaps]{Myriad Pro}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xltxtra}
\usepackage{multicol}
\usepackage[landscape]{geometry}
\usepackage[usenames,dvipsnames]{color}
\setcounter{secnumdepth}{-1}
\begin{document}
\setlength{\columnseprule}{1pt}
\setlength{\columnsep}{3em}
\section{Kwantlen Polytechnic University: Mission and Mandate}
\raggedcolumns
\begin{multicols}{2}


\subsection{Who We Are}
Kwantlen is British Columbia's Polytechnic University. As a leader in innovative and interdisciplinary education, we offer all learners, regardless of background and preparation, opportunities to achieve the highest standards of academic performance. Our many programs of study provide applied learning and broad-based university education.

Our relevant, engaging, and inclusive environment facilitates personal and professional development. Critical awareness, cultural sensitivity, and social responsibility are all foundational to our commitment. We provide accessible and outstanding scholarship to prepare our learners for a complex world.

We value scholarship as a socially relevant obligation and opportunity. We support multiple approaches to research and innovation to address community, industry, and market needs. Community engagement is the manner in which we demonstrate our commitment to the social value of the University.

Our approach to teaching and learning is based on mentorship, discovery, purposefulness, and learner autonomy. Our faculty demonstrate an authentic spirit of inquiry as a model for learners. In turn, learners engage with a teaching environment that is collaborative, innovative, creative, and respectful.

Our University culture is based on critical inquiry, collegial debate, knowledge generation, freedom of expression, diversity, and environmental stewardship and sustainability. Yet we go farther: by recognizing the necessary integration of these principles as our foundation and our purpose.

\columnbreak

\subsection{What We Do}

We support a dynamic educational community that embraces emerging and experimental teaching methods and research-driven pedagogy. We create a culture that is learner-focused, academically rigorous, innovative, interdisciplinary, and socially responsible.

We respect the academic freedom of learners, faculty, and staff as foundational to the pursuit of knowledge. We recognize that all members of the University community are responsible for exercising their freedoms with  honesty, integrity, and respect for the principles of justice, equality, and human rights.

\subsubsection{\color{Maroon}Access}
Kwantlen's history and purpose as a polytechnic institution integrates college, trades, and university experiences. We respond to community, regional, industry, and market needs through a distinctive variety of programs, research initiatives, and community partnerships. Transitions programs, multiple entry points, international education, and bridging opportunities demonstrate our commitment to accessible lifelong learning across a broad range of educational options. We view access to education as the pathway to citizenship.

\subsubsection{\color{Maroon}Scholarship}
We encourage faculty and learners to participate in many forms of knowledge generation and research, including those focused on discovery, creativity, application, and teaching. We honour and reward scholarship that involves learners and the broader community in research design, development, and dissemination.

\subsubsection{\color{Maroon}Community}
We recognize and respond to challenges and opportunities within the dynamic communities of the Fraser region of British Columbia. We promote leadership by developing and supporting community-based initiatives that address economic, social, and cultural issues within an increasingly global context. Synergies among various groups are developed and sustained through complementary cultures and practices that encourage autonomy, adaptability, collegiality, service, and excellence.

\subsubsection{\color{Maroon}Mentorship}
We provide and promote a learning environment in which learners examine and develop their values, goals, and character through the integration of personal, academic, and professional inquiry. We encourage lifelong learning and a continuing relationship with the University. We expose learners to workplace experiences, offer programs in continuing education, and provide  educational opportunities for those seeking career change or development. 


\subsubsection{\color{Maroon}Stewardship}
Through program delivery, services, and research initiatives, we aspire to inform and transform attitudes and values to reflect our role as a responsible and sustainable educational institution. Our health and our achievements derive from, and depend upon, the health of the Earth and its inhabitants.

\vspace{1cm}
\noindent
\textsf{Final, January 2010}

\end{multicols}


\end{document}
