   \TeXXeTstate=1
\documentclass[letterpaper,10pt,DIV=9,headsepline]{scrreprt}
   \usepackage{fontspec,xunicode}
    \defaultfontfeatures{Numbers=OldStyle,Scale=MatchLowercase,Mapping=tex-text}
    \setmainfont{Adobe Jenson Pro}
    \setromanfont[Mapping=tex-text]{Adobe Jenson Pro}
    \setsansfont[Scale=MatchUppercase]{Brioso Pro}
   \usepackage{graphicx}
   \usepackage{xltxtra}

\usepackage{scrpage2}
\clubpenalty=6000
\widowpenalty=6000
\author{Ross Laird}
\title{IDEA 3100 Evaluation for Assignment 2}
\date{today}
\ohead{IDEA 3100}
\begin{document}
\pagestyle{scrheadings}
\setcounter{secnumdepth}{-1}

\subsubsection{Evaluation for Assignment 3}

Student: Edmond Au

\bigskip
Hi Edmond;

An excellent project, with clear and strong effort and engagement. I
particularly appreciated the fusion of emotion, imagery, craft, and
creative play of all sorts.

Nicely done.

Also, this seems like a very honest and personally reflective project.
Much vulnerability and self-awareness here.

%Nice job! Very engaging and interdisciplinary. And it sounds, as I
%listen to this, that you are enjoying yourself in these songs (and
%that you are stretching yourself, as well). That's good. I also like
%the resonant aspects of the music. Well done.

%I enjoyed this project. It seems simple on the surface, but it leads
%to some very interesting places. I enjoyed the dimensionality of it,
%the blending of actual and fictional, and the many connections and
%surprises to be found through your narrative.

\bigskip
Proposed evaluation: 18/20

\bigskip
As you know, one of the things we are trying to do in this course is
to challenge the essentially arbitrary structures and methods of
traditional academic work; this applies to evaluations as well.
Accordingly, I have not simply assigned a number to your evaluation
(well, I have assigned a number, but it's not necessarily the final
number). What I'd like you to do is review the criteria below, in
light of your knowledge of the extent to which you applied
interdisciplinary and creative strategies to this project. Do you
think that my evaluation is correct, too high, or too low? It's
important that you play a role in the evaluation, so please take some
time to think about this.

\begin{itemize}
\item Demonstration of commitment to interdisciplinary process.
\item Demonstration of commitment to the development of self-awareness.
\item Demonstration of commitment to the development of creativity.
\item Ability to participate in appropriate creative exploration.
\item Willingness to take appropriate risks and to challenge oneself.
\item Ability to take personal responsibility for learning.
\end{itemize}

If you think that my proposed number is correct -- that it accurately
reflects your engagement and effort for this project -- then we're in
complete agreement and you need do nothing. If you think that my
proposed number is either too high or too low, let me know and we'll
talk about it.


\end{document}
%%% Local Variables: %%% mode: latex %%% TeX-master: t %%% End:
